---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# naicsmatch

<!-- badges: start -->
<!-- badges: end -->

The goal of naicsmatch is to provide functions and crosswalks from and between NAICS sector codes

The North American Industry Classification system is the standard used by Federal statistical agencies in classifying businesses for the purposes of data collection, analysis, and publishing statistical data. [see here](https://www.census.gov/eos/www/naics/index.html)

The NAICS code classification system is regularly updated, and so translating between different iterations can be necessary to do comparisons between different data sources. This package provides some functions to automate some of these translations using concordances downloaded from the U.S. Census and other agencies and data sources.

References:

* [NAICS Concordances](https://www.census.gov/eos/www/naics/concordances/concordances.html)


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jameelalsalam/naicsmatch")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(tidyverse)
library(naicsmatch)

sum(ex_asm$vos)

ex_asm
## basic example code
```


```{r}

library(tidyverse)

naics_2007_2012

```


```{r}
asm_2012 <- ex_asm %>% left_join(naics_2007_2012, 
                     by = c("NAICS.id" = "naics_2007"))

sum(asm_2012$vos)

asm_2012
```


# Weighting

```{r}
naics_2007_2012_wgt <- naics_2007_2012 %>%
  group_by(naics_2007) %>%
  summarize(wgt = 1 / n())

filter(naics_2007_2012_wgt, wgt != 1)
```


```{r}
asm_2012_v2 <- asm_2012 %>%
  left_join(naics_2007_2012_wgt, 
            by = c("NAICS.id" = "naics_2007")) %>%
  mutate(vos = vos * wgt)

sum(asm_2012_v2$vos)

asm_2012_v2
```

# Aggregation

```{r}

asm_2012_v3 <- asm_2012_v2 %>%
  group_by(naics_2012) %>%
  summarize(vos = sum(vos, na.rm = TRUE))

sum(asm_2012_v3$vos)

asm_2012_v3
```


